%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

% \begin{flushright}{\slshape
%     We have seen that computer programming is an art, \\
%     because it applies accumulated knowledge to the world, \\
%     because it requires skill and ingenuity, and especially \\
%     because it produces objects of beauty.} \\ \medskip
%     --- \defcitealias{knuth:1974}{Donald E. Knuth}\citetalias{knuth:1974} \citep{knuth:1974}
% \end{flushright}

% Where the research activity for the thesis was undertaken jointly with others,
% the name of such collaborators or co-authors must be listed immediately following the
% statement of original authorship page, including a short description of the nature of the
% contribution made by each author, including the candidate

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}
Firstly, I want to express my greatest appreciation to my supervising professor, Prof. Donal Finn, for his invaluable guidance and feedback throughout the research project. His input and encouragement have been essential in shaping the direction of this project.

I also extend my gratitude to the project collaborator, Dr Mohammad Saffari, for his contributions to this work. His expertise in \texttt{EnergyPlus} (and accompanying software) and willingness to share data to carry out the analysis has been instrumental in the success of this project, and I am thankful for his enthusiasm to collaborate and share his knowledge.

I want to thank Dr Alessandro Maccarini for His Modelica/Dymola course in Copenhagen, which provided me with a solid foundation in the subject matter. His dedication to teaching has been inspiring. Thank you to UCD for providing funding to visit Copenhagen, and software licenses. Thank you to Dr Michael Wetter and the Modelica Buildings Library team at Lawrence Berkeley National Laboratory for their work on progressing the building energy simulation field. 

I cannot forget to acknowledge the unwavering support and patience of my family and friends, who have been a constant source of encouragement throughout my academic journey. Thank you to Linus Sebastian, Michael Stevens, Derek Muller, Hank Green, and my late grandfather J\"urgen Bielstein for being the people for ultimately influenced me to pursue mechanical engineering. 


Special thanks to fellow Modelica-enjoyer Matthew Duffy, my sauna-buddy Daniel Joyce and Cormac Moloney for being my debugging rubber duck. 

Finally, thank you to my parents for giving me the opportunity to attend university, providing for me and always wanting the best for me. Thank you, I love you.



% Many thanks to everybody who already sent me a postcard!

% Regarding the typography and other help, many thanks go to Marco
% Kuhlmann, Philipp Lehman, Lothar Schlesier, Jim Young, Lorenzo
% Pantieri and Enrico Gregorio\footnote{Members of GuIT (Gruppo
% Italiano Utilizzatori di \TeX\ e \LaTeX )}, J\"org Sommer,
% Joachim K\"ostler, Daniel Gottschlag, Denis Aydin, Paride
% Legovini, Steffen Prochnow, Nicolas Repp, Hinrich Harms,
% Roland Winkler, Jörg Weber, Henri Menke, Claus Lahiri,
% Clemens Niederberger, Stefano Bragaglia, Jörn Hees,
% Scott Lowe, Dave Howcroft, Jos\'e M. Alcaide, David Carlisle,
% Ulrike Fischer, Hugues de Lassus, Csaba Hajdu, Dave Howcroft, 
% and the whole \LaTeX-community for support, ideas and
% some great software.

% \bigskip

% \noindent\emph{Regarding \mLyX}: The \mLyX\ port was intially done by
% \emph{Nicholas Mariette} in March 2009 and continued by
% \emph{Ivo Pletikosi\'c} in 2011. Thank you very much for your
% work and for the contributions to the original style.


\endgroup
